---
title: "Sc analysis"
author: "Luisa Santus"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE,results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(Matrix)
library(scater)
library(SingleCellExperiment)
library(purrr)
library(stringr)
library(mvoutlier)
library(dropbead)
library(scater)
library(stringr)
library(scran)
library(cowplot)
library(rtracklayer)
library(org.Mmu.eg.db)
library(graphics)
library(RCy3)
library(ComplexHeatmap)
library(org.Hs.eg.db)

source("../utils/02_sc_utils.R")
source("../utils/03_de_utils.R")

ref <- import(file.path("/home/luisas/Desktop/cluster/data/01_bulk_RNA-Seq_lncRNAs_annotation/old/03_novel_lncrnas/02_final_catalogue/rheMac10_EBOV-Kikivit_and_novelcatalogue_with_names.gtf"))
ebola_ref <- import(file.path("/home/luisas/Desktop/cluster/data/00_RawData/pardis_shared_data/sabeti-txnomics/shared-resources/HISAT2/EBOV-Kikwit/KU182905.1.gtf"))
result_path <- file.path("/home/luisas/Desktop/cluster/data/RObjects/results")
robjectsdir<- "/home/luisas/Desktop/cluster/data/RObjects_old/"

## Here add path 
immune.combined <- readRDS("/home/luisas/Desktop/cluster/data/RObjects_old/invivo_immunecombined_cellqc_aftercellandgeneqc_afterScrublet_integrated.rds")
```


# ------------------------------------
# Differential expression analysis
# ------------------------------------

# Group DPI to gather more 
```{r cars}
immune.combined$group <- immune.combined$dpi

immune.combined$group <- gsub("D-04" , "baseline",immune.combined$group)
immune.combined$group <- gsub("D000" , "baseline",immune.combined$group)

immune.combined$group <- gsub("D003" , "early",immune.combined$group)

immune.combined$group <- gsub("D004" , "middle",immune.combined$group)
immune.combined$group <- gsub("D005" , "middle",immune.combined$group)

immune.combined$group <- gsub("D006" , "late",immune.combined$group)
immune.combined$group <- gsub("D007" , "late",immune.combined$group)
immune.combined$group <- gsub("D008" , "late",immune.combined$group)

table(immune.combined$group)
DimPlot(immune.combined, group.by = "group", reduction = "umap")

#saveRDS(immune.combined, file.path(robjectsdir, "immune.combined_invivo_aftercomplete_preprocessing.rds"))
```
# -----------------------------------
# Specificity Matrix
# -----------------------------------
```{r cars}
library(dplyr)
library(ComplexHeatmap)
#immune.combined <- readRDS(file.path(robjectsdir, "immune.combined_invivo_aftercomplete_preprocessing.rds"))
pc <- subset(immune.combined, features = annotated_mrnas)
lnc <- subset(immune.combined, features = all_lncrnas)

calc_z <- function(expression){
  mean <- rowMeans(expression); sd <- apply(X = expression,MARGIN = 1 , FUN = sd); 
  z <- (expression-mean)/sd
  colnames(z) <- colnames(expression)
  return(z)
}

get_z_scores <- function(seuratobject){
  # First computes z score per gene per cell and then calculate the mean of z score per celltype
  expression<- as.matrix(seuratobject@assays$RNA@data)
  colnames(expression) <- Idents(seuratobject)
  zmatrix<- calc_z(expression)
  averaged_z_matrix <- sapply(unique(colnames(zmatrix)), function(x) rowMeans(zmatrix[,colnames(zmatrix) ==x]))
  return(averaged_z_matrix)
}


pc_z_m <- get_z_scores(pc)
lnc_z_m <- get_z_scores(lnc)

column_ha <- HeatmapAnnotation(celltype = c("B","Myeloid", "T"))
hm_pc <-  Heatmap(pc_z_m[rowSums(is.na(pc_z_m)) ==0,], show_row_names = FALSE, top_annotation = column_ha, column_order = c("B","Myeloid", "T")) 
hm_lnc <-  Heatmap(lnc_z_m[rowSums(is.na(lnc_z_m)) ==0,], show_row_names = FALSE, top_annotation = column_ha, column_order = c("B","Myeloid", "T")) 


hm_pc
hm_lnc


# ------------- second option 
# now calculate the medians and then the z score 
expression_pc <- as.matrix(pc@assays$RNA@data)
colnames(expression_pc) <- Idents(pc)
expression_lnc <- as.matrix(lnc@assays$RNA@data)
colnames(expression_lnc) <- Idents(lnc)

means <- sapply(unique(colnames(expression_pc)), function(x) rowMeans(expression_pc[,colnames(expression_pc) ==x]))
rownames(means) <- rownames(expression_pc)
zm_medians <- calc_z(means)
Heatmap(zm_medians[rowSums(is.na(zm_medians)) ==0,], show_row_names = FALSE) 


means <- sapply(unique(colnames(expression_lnc)), function(x) rowMeans(expression_lnc[,colnames(expression_lnc) ==x]))
rownames(means) <- rownames(expression_lnc)
zm_medians <- calc_z(means)
Heatmap(zm_medians[rowSums(is.na(zm_medians)) ==0,], show_row_names = FALSE) 
unique(Idents(pc))
pc <- subset(pc, idenst = "Myeloid")
expression_pc <- as.matrix(pc@assays$RNA@data)
z_scores_lnc <- calc_z(expression_pc)
test <- z_scores_lnc[,1:1000]
Heatmap(test, show_row_names = FALSE, show_column_names = FALSE, cluster_rows = TRUE)

PCAPlot(immune.combined)
PCAPlot(pc)
pc <- RunPCA(pc)
lnc <- RunPCA(lnc)
PCAPlot(lnc)

immune.combined$group

```





# -----------------------------------
# Differential Expression Analysis 
# -----------------------------------

```{r cars}


mrna_ref_id <- unique(ref$gene_name[ref$gene_biotype == "protein_coding"])
#immune.combined <- readRDS( file.path(robjectsdir, "immune.combined_invivo_aftercomplete_preprocessing.rds"))
annotated_mrnas <- rownames(immune.combined)[rownames(immune.combined) %in% str_replace_all(mrna_ref_id, "_" , "-")]

# ------------------------------------
#     LncRNAs present in object 
# -------------------------------------
# if i want to add also mirnas
#lnc_ref_id <- unique(ref$gene_name[ref$gene_biotype %in% c("lncRNA", "miRNA")])
lnc_ref_id <- unique(ref$gene_name[ref$gene_biotype %in% c("lncRNA")])


# i remove the gene names so i can match 
annotation_with_no_genename <- lapply(lnc_ref_id, function(x) str_split(x, "_")[[1]][1])
annotated_lncrnas <- rownames(immune.combined)[ lapply(rownames(immune.combined), function(x) str_split(x, "-")[[1]][1]) %in% annotation_with_no_genename]
annotated_lncrna_gene_names_only <- lnc_ref_id[!grepl("_unkown",lnc_ref_id)] [lnc_ref_id[!grepl("_unkown",lnc_ref_id)] %in%  rownames(immune.combined)]
complete_set_annotated_lncrnas_in_object <- c(annotated_lncrnas, annotated_lncrna_gene_names_only)
length(unique(complete_set_annotated_lncrnas_in_object))

# Novel lncRNAs
ribodepl_genes <- rownames(immune.combined)[grepl("ribodepl*", rownames(immune.combined))]
polya_genes <- rownames(immune.combined)[grepl("polya*", rownames(immune.combined))]
novel_lncrnas <- c(ribodepl_genes, polya_genes)

# Complete Set 
all_lncrnas <- c(complete_set_annotated_lncrnas_in_object, novel_lncrnas)
all_lncrnas <- all_lncrnas[!grepl("ribodepl-MSTRG.72510-unkown",all_lncrnas)]
```


```{r cars}
library(MAST)
get_mast_model  <- function(subset, relevel = "bystander", contarst = "conditioninfected"){
    # Convert to single cell experiment
    subset <- as.SingleCellExperiment(subset)
    # Extract ident column and use it as level
    cond<-factor(colData(subset)$ident)
    cond<-relevel(cond,relevel)
    colData(subset)$cond<-cond
    colData(subset)$wellKey<-colnames(subset)
    # add individual column in case is missing
    
    scam <- as(subset, 'SingleCellAssay')
    
    rowData(scam)$primerid <- rownames(rowData(scam))
   
    cdr2 <-colSums(assay(scam)>0)
     
    colData(scam)$cngeneson <- scale(cdr2)
    
    #We’ll fit a hurdle model, modeling the condition and (centered) ngeneson factor, thus adjusting for the cellular detection rate.
    #In order to have more interpretable coefficients, we’ll set the reference level of the factor to be the “unstimulated” cells.
    cond<-factor(colData(scam)$cond)
    cond<-relevel(cond,relevel)
    colData(scam)$condition<-cond
    
    zlmCond <- zlm(~condition + cngeneson+ freshfrozen , scam)
    # Summarize data
    print("here")
    summaryCond <- summary(zlmCond, doLRT=contarst) 
    summaryDt <- summaryCond$datatable
    fcHurdle <- merge(summaryDt[contrast==contarst & component=='H',.(primerid, `Pr(>Chisq)`)], #hurdle P values
                          summaryDt[contrast==contarst & component=='logFC', .(primerid, coef, ci.hi, ci.lo)], by='primerid') #logFC coefficients
    fcHurdle[,fdr:=p.adjust(`Pr(>Chisq)`, 'fdr')]
    return(fcHurdle)
}
```
# Set up differential expression analysis
```{r cars}
immune.combined$freshfrozen <- as.factor(ifelse(immune.combined$batch =="fresh", 0, 1))

myeloids <- subset(immune.combined, idents = "Myeloid"); Idents(myeloids) <- myeloids$group
#nk <- subset(immune.combined, idents = "NK"); Idents(nk) <- nk$group
t <- subset(immune.combined, idents = "T"); Idents(t) <- t$group
b <- subset(immune.combined, idents = "B"); Idents(b) <- b$group


mast_stateVSbaseline_perCelltype <- function(immune.combined , celltype, idents,ebola_genes = ebola_genes){
  subset <- subset(immune.combined, idents = celltype); Idents(subset) <- subset$group
  subset <- subset(subset, idents = idents)
  #Remove ebola genes
  
  genes_no_ebola <-rownames(subset)[!(rownames(subset) %in% ebola_genes)]
  subset <- subset(subset, features = genes_no_ebola)
  
  model_late_subset <- get_mast_model(subset, relevel = idents[1], contarst = paste0("condition", idents[2]))
  saveRDS(model_late_subset,file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_", idents[2],".rds")))
  
  return(model_late_subset)
}



mast_myeloid_late <- mast_stateVSbaseline_perCelltype(immune.combined = immune.combined, "Myeloid", c("baseline", "late"),ebola_genes)
mast_myeloid_middle <- mast_stateVSbaseline_perCelltype(immune.combined, "Myeloid", c("baseline", "middle"),ebola_genes)
mast_myeloid_early <- mast_stateVSbaseline_perCelltype(immune.combined, "Myeloid", c("baseline", "early"), ebola_genes)

mast_T_late <- mast_stateVSbaseline_perCelltype(immune.combined, "T", c("baseline", "late"), ebola_genes)
mast_T_middle <- mast_stateVSbaseline_perCelltype(immune.combined, "T", c("baseline", "middle"), ebola_genes)
mast_T_early <- mast_stateVSbaseline_perCelltype(immune.combined, "T", c("baseline", "early"), ebola_genes)

mast_B_late <- mast_stateVSbaseline_perCelltype(immune.combined, "B", c("baseline", "late"), ebola_genes)
mast_B_middle <- mast_stateVSbaseline_perCelltype(immune.combined, "B", c("baseline", "middle"), ebola_genes)
mast_B_early <- mast_stateVSbaseline_perCelltype(immune.combined, "B", c("baseline", "early"), ebola_genes)



celltype <- "Myeloid"
mast_myeloid_late <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","late",".rds")))
mast_myeloid_middle <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","middle",".rds")))
mast_myeloid_early <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","early",".rds")))

celltype <- "B"
mast_B_late <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","late",".rds")))
mast_B_middle <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","middle",".rds")))
mast_B_early <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","early",".rds")))

celltype <- "T"
mast_T_late <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","late",".rds")))
mast_T_middle <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","middle",".rds")))
mast_T_early <- readRDS(file.path(robjectsdir, paste0("MAST_invivo_model_", celltype,"_","early",".rds")))


# Missing Global one 
```

```{r heatmap}
hm <- function(m,features, row_title, first  = FALSE, row_names = FALSE){
  # Only select lncrnas
  m <- m[rownames(m) %in% features,, drop = FALSE]
    rownames(m) <- gsub("-unkown","",rownames(m))


      column_ha <- HeatmapAnnotation(celltype = unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[1])), 
                        comparison = unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[2])),
                        show_annotation_name = FALSE
                        )
  row_ha <- rowAnnotation(genetype =substring(rownames(m), 1,3),
                        show_annotation_name = FALSE)
  if(!first){

    h1 <- Heatmap(m, show_row_dend = FALSE,
          show_column_dend = FALSE,
          cluster_columns = FALSE,
          show_column_names = FALSE,
          show_row_names = row_names, 
          row_title = row_title,
          row_title_rot = 0, 
          column_split = unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[1])), 
          show_heatmap_legend = FALSE
          )    
  }else{

    h1 <- Heatmap(m, show_row_dend = FALSE,
          show_column_dend = FALSE,
          cluster_columns = FALSE,
          show_column_names = FALSE,
          show_row_names = row_names, 
          row_title = row_title,
          row_title_rot = 0, 
          top_annotation = column_ha, 
          
          column_split = unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[1]))
          )
  }
  print("done")
  return(h1)
}
```

# ---------------------------------
#       LNCRNAS
# ---------------------------------

```{r heatmap}

set.seed(123)

# -------------------------------------------------------------------------
# Gather all information about changes of each gene in each comparison per celltype
# -------------------------------------------------------------------------

# ------- Myeloid
celltype <- "myeloid"
mast_myeloid_late$comparison <- paste0(celltype,"_late")
mast_myeloid_middle$comparison <- paste0(celltype,"_middle")
mast_myeloid_early$comparison <- paste0(celltype,"_early")
myeloids_de <- rbind(mast_myeloid_late, mast_myeloid_middle, mast_myeloid_early)

# ------- T
celltype <- "T"
mast_T_late$comparison <- paste0(celltype,"_late")
mast_T_middle$comparison <- paste0(celltype,"_middle")
mast_T_early$comparison <- paste0(celltype,"_early")
t_de <- rbind(mast_T_late, mast_T_middle, mast_T_early)


# ------- B
celltype <- "B"
mast_B_late$comparison <- paste0(celltype,"_late")
mast_B_middle$comparison <- paste0(celltype,"_middle")
mast_B_early$comparison <- paste0(celltype,"_early")
b_de <- rbind(mast_B_late, mast_B_middle, mast_B_early)

stats_complete <- rbind( myeloids_de, t_de, b_de)

# Convert into matrix
m <-xtabs(coef~primerid+comparison, stats_complete)
attr(m, "class") <- NULL
colnames(m)
m <- m[,c("myeloid_early" ,"myeloid_middle","myeloid_late", "B_early" ,"B_middle","B_late", "T_early" ,"T_middle","T_late")]

m_significance<-xtabs(fdr~primerid+comparison, stats_complete)
attr(m_significance, "class") <- NULL
m_significance <- m_significance[,c("myeloid_early" ,"myeloid_middle","myeloid_late", "B_early" ,"B_middle","B_late", "T_early" ,"T_middle","T_late")]

# -------------------------------------------------------------------------
# Select genes
# -------------------------------------------------------------------------

FDRTHRESHOLD <- 0.1

get_de_names <- function(df_filtered, FDRTHRESHOLD = 0.1, FCTHRESHOLD = log(1.2), subset = all_lncrnas){
  # Only select lncrnas 
  df_filtered <- df_filtered[df_filtered$primerid %in% subset,]
  # Only select significant ones
  df_filtered <- df_filtered[fdr<FDRTHRESHOLD &abs(coef) > FCTHRESHOLD,]
  return(df_filtered$primerid)
}


# Select genes to plot
lnc_myeloid_late <- get_de_names(mast_myeloid_late, subset = all_lncrnas)
#intersect(bulk_de_genes_all, gsub("-", "_",gsub("-unkown", "", lnc_myeloid_late)))

lnc_myeloid_middle <- get_de_names(mast_myeloid_middle, subset = all_lncrnas)
lnc_myeloid_early <- get_de_names(mast_myeloid_early, subset = all_lncrnas)

lnc_T_late <- get_de_names(mast_T_late, subset = all_lncrnas)
lnc_T_middle <- get_de_names(mast_T_middle, subset = all_lncrnas)
lnc_T_early <- get_de_names(mast_T_early, subset = all_lncrnas)

lnc_B_late <- get_de_names(mast_B_late, subset = all_lncrnas)
lnc_B_middle <- get_de_names(mast_B_middle, subset = all_lncrnas)
lnc_B_early <- get_de_names(mast_B_early, subset = all_lncrnas)

## mrnas
mrna_myeloid_late <- get_de_names(mast_myeloid_late, subset = annotated_mrnas)
mrna_myeloid_middle <- get_de_names(mast_myeloid_middle, subset = annotated_mrnas)
mrna_myeloid_early <- get_de_names(mast_myeloid_early, subset = annotated_mrnas)

mrna_T_late <- get_de_names(mast_T_late, subset = annotated_mrnas)
mrna_T_middle <- get_de_names(mast_T_middle, subset = annotated_mrnas)
mrna_T_early <- get_de_names(mast_T_early, subset = annotated_mrnas)

mrna_B_late <- get_de_names(mast_B_late, subset = annotated_mrnas)
mrna_B_middle <- get_de_names(mast_B_middle, subset = annotated_mrnas)
mrna_B_early <- get_de_names(mast_B_early, subset = annotated_mrnas)
# ----------------------------


de_lists_lnc <- list(lnc_myeloid_late, lnc_myeloid_middle, lnc_myeloid_early,lnc_T_late, lnc_T_middle, lnc_T_early, lnc_B_late, lnc_B_middle, lnc_B_early )
names(de_lists_lnc) <- c("lnc_myeloid_late", "lnc_myeloid_middle", "lnc_myeloid_early","lnc_T_late", "lnc_T_middle", "lnc_T_early", "lnc_B_late", "lnc_B_middle", "lnc_B_early")


de_lists_mrna <- list(mrna_myeloid_late, mrna_myeloid_middle, mrna_myeloid_early,mrna_T_late, mrna_T_middle, mrna_T_early, mrna_B_late, mrna_B_middle, mrna_B_early )
names(de_lists_mrna) <- c("mrna_myeloid_late", "mrna_myeloid_middle", "mrna_myeloid_early","mrna_T_late", "mrna_T_middle", "mrna_T_early", "mrna_B_late", "mrna_B_middle", "mrna_B_early")


# -------------------------------------------------------------------------
# Create Heatmap(s)
# -------------------------------------------------------------------------
de_lists_lnc <- de_lists_lnc[unlist(lapply(1:length(de_lists_lnc), function(index) length(de_lists_lnc[[index]]) != 0))]
h1_lnc<- hm(m= m,row_title = names(de_lists_lnc)[1], features  = de_lists_lnc[1][[1]], first = TRUE, row_names = TRUE)
hms <- lapply(2:length(de_lists_lnc), function(index) hm(m,row_title = names(de_lists_lnc)[index], features  = de_lists_lnc[index][[1]], row_names = TRUE))

length(hms)
heatmap_lnc <- h1_lnc %v% hms[[1]]%v%hms[[2]]%v%hms[[3]]%v%hms[[4]]%v%hms[[5]]%v%hms[[6]]
heatmap_lnc

# mrna ---------
de_lists_mrna <- de_lists_mrna[unlist(lapply(1:length(de_lists_mrna), function(index) length(de_lists_mrna[[index]]) != 0))]
h1 <- hm(m= m,row_title = names(de_lists_mrna)[1], features  = de_lists_mrna[1][[1]], first = TRUE, row_names = FALSE)
hms_mrna <- lapply(2:length(de_lists_mrna), function(index) hm(m,row_title = names(de_lists_mrna)[index], features  = de_lists_mrna[index][[1]], row_names = FALSE))

length(de_lists_mrna)

heatmap_mrna <- h1 %v% hms_mrna[[1]]%v%hms_mrna[[2]]%v%hms_mrna[[3]]%v%hms_mrna[[4]]%v%hms_mrna[[5]]%v%hms_mrna[[6]]%v%hms_mrna[[7]]%v%hms_mrna[[8]]
heatmap_mrna

do_go(unlist(de_lists_mrna))
```

# ------------------------------
#       Different heatmaps 
# ------------------------------

```{r a}
FDRTHRESHOLD <- 0.1
FCTHRESHOLD <- log(1.2)
# Obtain Genes To Plot 
lnc_myeloid_late <- get_de_names(mast_myeloid_late, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
lnc_myeloid_middle <- get_de_names(mast_myeloid_middle, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
lnc_myeloid_early <- get_de_names(mast_myeloid_early, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
de_lists_lnc_my <- list(lnc_myeloid_late, lnc_myeloid_middle, lnc_myeloid_early )
names(de_lists_lnc_my) <- c("lnc_myeloid_late", "lnc_myeloid_middle", "lnc_myeloid_early")



lnc_T_late <- get_de_names(mast_T_late, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
lnc_T_middle <- get_de_names(mast_T_middle, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
lnc_T_early <- get_de_names(mast_T_early, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
de_lists_lnc_T <- list(lnc_T_late, lnc_T_middle, lnc_T_early )
names(de_lists_lnc_T) <- c("lnc_T_late", "lnc_T_middle", "lnc_T_early")
de_lists_lnc_T <- de_lists_lnc_T[unlist(lapply(1:length(de_lists_lnc_T), function(index) length(de_lists_lnc_T[[index]]) != 0))]

lnc_B_late <- get_de_names(mast_B_late, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
lnc_B_middle <- get_de_names(mast_B_middle, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
lnc_B_early <- get_de_names(mast_B_early, subset = all_lncrnas, FDRTHRESHOLD = FDRTHRESHOLD, FCTHRESHOLD = FCTHRESHOLD)
de_lists_lnc_B <- list(lnc_B_late, lnc_B_middle, lnc_B_early )

names(de_lists_lnc_B) <- c("lnc_B_late", "lnc_B_middle", "lnc_B_early")
de_lists_lnc_B <- de_lists_lnc_B[unlist(lapply(1:length(de_lists_lnc_B), function(index) length(de_lists_lnc_B[[index]]) != 0))]

de_lnc_all <- c(de_lists_lnc_my, de_lists_lnc_T, de_lists_lnc_B)

# Construct matrix
# Convert into matrix
m <-xtabs(coef~primerid+comparison, stats_complete)
attr(m, "class") <- NULL
m <- m[,c("myeloid_early" ,"myeloid_middle","myeloid_late", "B_early" ,"B_middle","B_late", "T_early" ,"T_middle","T_late")]
m <- m[rownames(m) %in% unique(unlist(de_lnc_all)),, drop = FALSE]
#m <- m[, c("myeloid_early" ,"myeloid_middle","myeloid_late")]
rownames(m) <- gsub("-unkown","",rownames(m))

m_fdr <-xtabs(fdr~primerid+comparison, stats_complete)
attr(m_fdr, "class") <- NULL
m_fdr <- m_fdr[,c("myeloid_early" ,"myeloid_middle","myeloid_late", "B_early" ,"B_middle","B_late", "T_early" ,"T_middle","T_late")]
m_fdr <- m_fdr[rownames(m_fdr) %in% unique(unlist(de_lnc_all)),, drop = FALSE]
#m_fdr <- m_fdr[, c("myeloid_early" ,"myeloid_middle","myeloid_late")]
rownames(m_fdr) <- gsub("-unkown","",rownames(m_fdr))

column_ha <- HeatmapAnnotation(celltype = unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[1])), 
                        comparison = unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[2])),
                        show_annotation_name = FALSE, 
                        col = list(celltype = c("myeloid" = brewer.pal(3, "Set2")[3],
                                                "T" = brewer.pal(3, "Set2")[1],
                                                "B" = brewer.pal(3, "Set2")[2]),
                                   comparison = c("early" = brewer.pal(8, "Pastel2")[6],
                                                "middle" = brewer.pal(8, "Pastel2")[7],
                                                "late" = brewer.pal(8, "Pastel2")[8])), 
                      
                        show_legend = FALSE
                        )
row_ha <- rowAnnotation(genetype =substring(rownames(m), 1,3),
                        show_annotation_name = FALSE)

lex_order <- match( gsub("-unkown","",unique(unlist(de_lnc_all))),rownames(m))

leg <- Legend(pch = "*", type = "points", labels = "FDR < 0.1 & |logFC|>log(1.2)")
leg_comparison <- Legend(at = c("Early VS Baseline", "Middle VS Baseline", "Late VS Baseline"), title = "Comparison", legend_gp = gpar(fill = brewer.pal(8,"Pastel2")[6:9]))
h <- Heatmap(m,name = "logFC", show_row_dend = FALSE,
          show_column_dend = FALSE,
          cluster_columns = FALSE,
          cluster_rows = FALSE,
          row_order = lex_order,
          
          cluster_row_slices = FALSE,
          show_column_names = FALSE,
          show_row_names = FALSE, 
          row_title_rot = 0,
          top_annotation = column_ha, 
         
          column_split = gsub("myeloid", "Myeloid",unlist(lapply( colnames(m), function(x) unlist(str_split(x, "_"))[1]))),
          
          cell_fun = function(j, i, x, y, width, height, fill) {
              if(m_fdr[i, j] < FDRTHRESHOLD & abs(m[i,j]) > FCTHRESHOLD )
                    grid.text(sprintf("*"), x, y, gp = gpar(fontsize = 10))}
          )

draw(h, annotation_legend_list = c(leg_comparison, leg))
# ------------------------------------
# Now try with z scores 
# ------------------------------------
calc_z <- function(matrix){
  mean <- rowMeans(matrix);
  sd <- apply(matrix,1, sd)
  z_scores <-apply(matrix,2, function(x) (x-mean)/sd)
  return(z_scores)
}


B <- subset(immune.combined, idents = "B")
T <- subset(immune.combined, idents = "T")
Myeloids <- subset(immune.combined, idents = "Myeloid")


# identify if upregulated or downregulated 

celltype <- "B"
de_lnc_all <- de_lists_lnc_B

 
z_score_heatmap <- function(subset, celltype,de_lnc_all, title, color_title){
  # Exprssion matrix
  matrix <- as.matrix(subset@assays$RNA@data)
  subset$state <- gsub("Myeloid", "myeloid",paste(Idents(subset), subset$group, sep = "_"))
  colnames(matrix) <- subset$state
  # Calculate Z scores 
  mean_matrix <- sapply(unique(colnames(matrix)), function(x) rowMeans(matrix[,colnames(matrix) ==x, drop =FALSE],  na.rm = TRUE))
  mean_matrix <- mean_matrix[rownames(mean_matrix) %in% unique(unlist(de_lnc_all)),, drop = FALSE]
  z_score_mean_matrix <- calc_z(mean_matrix)
  
  
  
 
  
  # Create annotations
  rownames(z_score_mean_matrix) <- gsub("-unkown","",rownames(z_score_mean_matrix))
  colnames(z_score_mean_matrix) <- gsub("Myeloid", "myeloid", colnames(z_score_mean_matrix))
  column_ha <- HeatmapAnnotation( 
                          comparison = anno_text( unlist(lapply(unlist(lapply( colnames(z_score_mean_matrix), function(x) unlist(str_split(x, "_"))[2])), function(x) toupper(substr(x,1,1)))), rot = 0, gp = gpar(fontsize = 25, fill =  brewer.pal(8, "Pastel2")[5:8]), location = 0.5, just = "center"),
                          show_annotation_name = FALSE, 
                          gap = unit(16, "points"),
                          
                          col = list(celltype = c("myeloid" = brewer.pal(3, "Set2")[3],
                                                  "T" = brewer.pal(3, "Set2")[1],
                                                  "B" = brewer.pal(3, "Set2")[2]),
                                     comparison = c("baseline" = brewer.pal(8, "Pastel2")[5],"early" = brewer.pal(8, "Pastel2")[6],
                                                  "middle" = brewer.pal(8, "Pastel2")[7],
                                                  "late" = brewer.pal(8, "Pastel2")[8])), 
                        
                          show_legend = FALSE, 
                           annotation_height =  unit(c(1), "cm")
                          )
  
  row_ha <- rowAnnotation(genetype =substring(rownames(z_score_mean_matrix), 1,3),show_legend = FALSE,
                        show_annotation_name = FALSE,
                        col = list(genetype = c("ENS" = brewer.pal(3, "Paired")[3],
                                                  "rib" = brewer.pal(3, "Paired")[1],
                                                  "pol" = brewer.pal(3, "Paired")[2])))
  h <- Heatmap(z_score_mean_matrix,name = "Z-Score", show_row_dend = FALSE, heatmap_width = unit(18, "cm"), 
            show_column_dend = FALSE,
            heatmap_height = unit(30, "cm"),
            cluster_columns = FALSE,
            cluster_rows = TRUE,
            cluster_row_slices = FALSE,
            show_column_names = FALSE,
            show_row_names = TRUE,
            row_names_gp = gpar(fontsize = 25),
            row_title_rot = 0,
            right_annotation = row_ha, 
            top_annotation = column_ha, 
            column_title = title, 
            column_title_gp = gpar(fill = color_title, fontsize =40, border = NA), 
            show_heatmap_legend = FALSE
           
           )
  h
  # Add legend
  leg_comparison <- Legend(at = c("Baseline", "Early", "Middle", "Late"), title = "Stage", legend_gp = gpar(fill = brewer.pal(8,"Pastel2")[5:9]), size = unit(30, "mm"))
  leg_annot <- Legend(at = c("Novel (Ribodepleted)", "Novel (Poly-A)", "Annotated"), title = "GeneType", legend_gp = gpar(fill = brewer.pal(3,"Paired")[1:3]),size = unit(30, "mm"))

  hm_wlegend <- draw(h,annotation_legend_list = c(leg_comparison, leg_annot))
  return(h)
}

z_score_heatmap(Myeloids, "myeloid", de_lists_lnc_my, "Myeloid",brewer.pal(3, "Set2")[3] )
z_score_heatmap(T, "T", de_lists_lnc_T, "T",  brewer.pal(3, "Set2")[1])
z_score_heatmap(B, "B", de_lists_lnc_B, "B",  brewer.pal(3, "Set2")[2])



```

# Save the heatmaps 
```{r colocation}
png(filename = "/home/luisas/Desktop/cluster/data/mnew.png", width = 1300, height = 1100)
z_score_heatmap(Myeloids, "myeloid", de_lists_lnc_my, "Myeloid",brewer.pal(3, "Set2")[3] )
dev.off()
```
```{r colocation}
png(filename = "/home/luisas/Desktop/cluster/data/tnew.png", width = 1300, height = 1100)
z_score_heatmap(T, "T", de_lists_lnc_T, "T",  brewer.pal(3, "Set2")[1])

dev.off()
```

```{r colocation}
png(filename = "/home/luisas/Desktop/cluster/data/bnew.png", width = 1300, height = 1100)
z_score_heatmap(B, "B", de_lists_lnc_B, "B",  brewer.pal(3, "Set2")[2])
dev.off()
```


# Venn Diagram 
```{r colocation}
library(VennDiagram)

# -------------------------------------------
# Overlap of Differentially expressed genes 
# -------------------------------------------
venn.diagram(
  x = list( unlist(unique(de_lists_lnc_T)), unlist(unique(de_lists_lnc_B)),unlist(unique(de_lists_lnc_my))),
  category.names = c( "T DE lncRNAs " , "B DE lncRNAs ","Myeloid DE lncRNAs" ),
  filename = '/home/luisas/Desktop/cluster/data/venn_diagramm_celltype',
  output = TRUE ,
          imagetype="png" ,
          height = 580 , 
          width = 580 , 
          resolution = 300,
          compression = "lzw",
          lwd = 3,
          col=brewer.pal(3, "Set2"),
          fill = c(alpha(brewer.pal(3, "Set2")[1],0.1), alpha(brewer.pal(3, "Set2")[2],0.1), alpha(brewer.pal(3, "Set2")[3],0.1)),
          cex = 0.8,
          fontfamily = "times",
          cat.cex = 0.7,
          cat.default.pos = "outer",
          cat.pos = c(-165, 165, -0),
          cat.dist = c(0.085, 0.085, 0.085),
          cat.fontfamily = "times",
          cat.col = "black",
          rotation = 1
        )

# -------------------------------------------
# overlap stage specific genes
# -------------------------------------------
early_de <- c(de_lists_lnc_T$lnc_T_early, de_lists_lnc_B$lnc_B_early, de_lists_lnc_my$lnc_myeloid_early)
middle_de <- c(de_lists_lnc_T$lnc_T_middle, de_lists_lnc_B$lnc_B_middle, de_lists_lnc_my$lnc_myeloid_middle)
late_de <- c(de_lists_lnc_T$lnc_T_late, de_lists_lnc_B$lnc_B_late, de_lists_lnc_my$lnc_myeloid_late)

venn.diagram(
  x = list( unlist(unique(early_de)), unlist(unique(middle_de)),unlist(unique(late_de))),
  category.names = c( "Early DE lncRNAs" , "Middle DE lncRNAs ","Late DE lncRNAs" ),
  filename = '/home/luisas/Desktop/cluster/data/venn_diagram_stagessignature_de',
  output = TRUE ,
          imagetype="png" ,
          height = 680 , 
          width = 680 , 
          resolution = 300,
          compression = "lzw",
          lwd = 3,
          col=brewer.pal(8, "Pastel2")[6:8],
          
          cex = 0.8,
          fontfamily = "times",
          cat.cex = 0.7,
          cat.default.pos = "outer",
          cat.pos = c(-20, 20, -180),
          cat.dist = c(0.050, 0.050, 0.050),
          cat.fontfamily = "times",
          cat.col = "black",
          rotation = 1
        )

```


# -----------------------------------
#      Colocation Analysis
# ----------------------------------


```{r colocation}
# 1) For each gene lncrna find neighbour protein coding genes. ( either overlapping or in close region) --> putative pairs
# 2) Filter: Check the pearson correlation of the identified pairs, reatain the ones w/ |pearson correlation coefficient| > 0.6
# ==> Only if something nice comes out 
# 3) Validation: Simulate random pairs 
# 4) Validation: Compare the correlation of the ones identified with the random ones. 
library(zeallot)

# 1. extract the count matrix
#count_matrix <- log2(myeloids@assays$RNA@counts+.5)
#count_matrix <- myeloids@assays$RNA@counts
#rownames(count_matrix) <- gsub("-unkown","",rownames(count_matrix))
#exp_matrix <-myeloids@assays$integrated@data

# 2. 
mrnas_ranges <- ref[ref$gene_name %in% mrna_ref_id  ]
lnc_ranges <- ref[ref$gene_id %in% gsub("-", "_",gsub("-unkown", "",all_lncrnas))]

# Add missing gene entries (as novel lncrnas only prsent transcript and exon entries)
genes_lnc_ranges <- get_gene_only(lnc_ranges)$gene_id
missing_gene_lines <- unique(lnc_ranges$gene_id)[!(unique(lnc_ranges$gene_id) %in% genes_lnc_ranges)]
gr_list <- lapply(missing_gene_lines, function(x) get_gene_range(x, lnc_ranges))
missing_gene_ranges <- do.call(base::c,gr_list)
lnc_ranges_complete <-c(get_gene_only(lnc_ranges), missing_gene_ranges)



# --------------------------------------------------------------
#           Get 5 closest genes to each gene of interest
# --------------------------------------------------------------
# Find pairs under maximum overlap
get_n_closest_gr <- function(gr_gene, gr_hits =mrnas_ranges , n = 5, maxgap =1000000L ){
  p <- findOverlapPairs(gr_gene, get_gene_only(gr_hits), maxgap=maxgap, ignore.strand = TRUE)
  pairs_list <- zipup(p)
  print(gr_gene$gene_name)
  # Compute thedistances and select the first 5
  distances <- unlist(lapply(pairs_list, function(pair) GenomicRanges::distance(pair[1], pair[2], ignore.strand = TRUE) ))
  if(!is.null(distances)){
    n <- min(length(pairs_list), n)
    closest_5_pairs <-  pairs_list[order(distances)][1:n]
  }else{
    return(NULL)
  }
  return(closest_5_pairs)
}

de_sc <- gsub("-", "_",gsub("-unkown", "", c(unlist(unique(de_lists_lnc_T)), unlist(unique(de_lists_lnc_B)),unlist(unique(de_lists_lnc_my)))))

# Only select the DE interestin
lnc_ranges_complete_de <- lnc_ranges_complete[lnc_ranges_complete$gene_id %in% gsub("-", "_",gsub("-unkown", "",de_sc)), ]

# maximum 1Mb
all_lnc_1mb <- lapply(1:(length(lnc_ranges_complete_de)), function(index) unlist(get_n_closest_gr(lnc_ranges_complete_de[index], mrnas_ranges, n = Inf, maxgap = 1000000L))[c(rep(FALSE,1), TRUE)]$gene_name)
do_go(unlist(all_lnc_1mb))

logfc <- stats_complete[stats_complete$primerid %in% unlist(unique(all_lnc_1mb)),]$coef
 ego <- clusterProfiler::enrichGO(gene = unlist(unique(all_lnc_1mb)),OrgDb =org.Hs.eg.db, keyType = "SYMBOL",ont = "BP",universe =unique(ref$gene_name))
library(clusterProfiler) 
cnetplot(ego, foldChange = 2^logfc, colorEdge = TRUE, node_label = "gene")

barplot(ego)+theme(axis.text.y = element_text(size = 17))
dotplot(ego)+theme(axis.text.y = element_text(size = 17))

# 1Mb, 15 closest genes
all_lnc_1mb_max10 <- lapply(1:(length(lnc_ranges_complete_de)), function(index) unlist(get_n_closest_gr(lnc_ranges_complete_de[index], mrnas_ranges, n = 15, maxgap = 1000000L))[c(rep(FALSE,1), TRUE)]$gene_name)
do_go(unlist(all_lnc_1mb_max10))

```

# Correlation 

# Calculate correlation measure 

```{r colocation}
counts <- matrix(rpois(20*50, 100), 20, 50)
group <- sample(c("A", "B"), size = 20, replace = TRUE)
devtools::install_github("tpq/propr")
library(propr)

# Calculate counts matrix
count_matrix <- immune.combined@assays$RNA@counts



index <- 14

calc_corr <- function(index){
    gene_name <- str_replace( lnc_ranges_complete_de[index]$gene_name, "_", "-")
    gene_counts <- as.vector(count_matrix[gene_name,])
    FeaturePlot(immune.combined, features = c(gene_name), order = TRUE, pt.size  = 1.5)
    
    close_genes <- str_replace(unlist(get_n_closest_gr(lnc_ranges_complete_de[index], mrnas_ranges, n = Inf, maxgap = 1000000L))[c(rep(FALSE,1), TRUE)]$gene_name,"_", "-")
    
    expression_close_genes <- as.matrix(count_matrix[rownames(count_matrix) %in% close_genes,, drop = FALSE])
    dim(expression_close_genes)
    print(index)
    if(nrow(expression_close_genes) > 0){

          rhos <- unlist(lapply(1:nrow(expression_close_genes), function(x) cor.test(gene_counts, expression_close_genes[x,], method = "spearman")$estimate))
    pvals <- unlist(lapply(1:nrow(expression_close_genes), function(x) cor.test(gene_counts, expression_close_genes[x,], method = "spearman")$p.value))
    print("hey")
    df <- data.frame(lnc = gene_name , genename = rownames(expression_close_genes)[1:nrow(expression_close_genes)], rho = rhos , pvals = pvals)
    }else{
      df <- NULL
    }
    print("aa")
    return(df)
}

cor_table <- rbindlist((lapply(1:(length(lnc_ranges_complete_de$gene_id)), function(x) calc_corr(x))))
sig_cor_table <- cor_table[[1]][ cor_table[[1]]$pvals < 0.05, ]
sig_cor_table
do_go(sig_cor_table$genename)
```






```{r example}
library(scales)
# Overlap is quite low, maybe we can mention if we select a specific lncRNA
bulk_de_genes_all <-  readRDS(file.path(robjectsdir, "bulkDEgenes.rds"))
de_sc <- gsub("-", "_",gsub("-unkown", "", c(unlist(unique(de_lists_lnc_T)), unlist(unique(de_lists_lnc_B)),unlist(unique(de_lists_lnc_my)))))
intersect(de_sc, bulk_de_genes_all)

de_sc
setdiff(de_lists_lnc_my$lnc_myeloid_middle, de_lists_lnc_my$lnc_myeloid_late)


# One Cell Type Specific Myeloid
plot_seperate_features(immune.combined, gene  = c("ENSMMUG00000052585-unkown"), ident = "group")

# One Cell Type Specific Myeloid
lapply(de_lists_lnc_my$lnc_myeloid_late, function(gene) plot_seperate_features(immune.combined, gene  = c(gene), ident = "group", pt.size = 1))

plot_seperate_features(immune.combined, gene  = c("ribodepl-MSTRG.3272-unkown"), ident = "group")


DimPlot(immune.combined, label = TRUE, pt.size = 0.1, cols=  brewer.pal(3, "Set2"))+theme_minimal()+ theme(panel.background = element_rect(fill = "white", colour = "grey50"), panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(legend.text = element_text(size=18))

# One Cell Type Specific B
plot_seperate_features(immune.combined, gene  = c("ENSMMUG00000061948-unkown"), ident = "group", pt.size = 0.7)
plot_seperate_features(immune.combined, gene  = c("PVRL1"), ident = "group", pt.size = 1)
plot_seperate_features(immune.combined, gene  = c("THY1"), ident = "group", pt.size = 0.7,colors =c('#DCD9DE','navy'))+theme_minimal()+ theme(panel.background = element_rect(fill = "white", colour = "grey50"), panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(legend.text = element_text(size=18))
plot_seperate_features(immune.combined, gene  = c("CCDC153"), ident = "group", pt.size = 1)
plot_seperate_features(immune.combined, gene  = c("ABCG4"), ident = "group", pt.size = 1)

rownames(immune.combined)[grep("PVRL", rownames(immune.combined))]
unlist(all_lnc_1mb)[unlist(all_lnc_1mb) %in% rownames(immune.combined)]
unlist(all_lnc_1mb)["PVRL1" %in% rownames(immune.combined)]
mast_myeloid_late[grep("ENSMMUG00000061948", mast_myeloid_late$primerid),]
annotated_mrnas[grep("ENSMMUG00000061948-unkown", all_lncrnas)]


all_lncrnas[all_lncrnas == "ENSMMUG00000061948-unkown"]
all_lncrnas[grep("ENSMMUG000000619",all_lncrnas)]

immune.combined
table(rownames(immune.combined)=="ENSMMUG00000061948-unkown")
m[m[,1] =="ENSMMUG00000061948-unkown",]
m
lnc_ranges_complete_de <- lnc_ranges_complete[lnc_ranges_complete$gene_id %in% gsub("-", "_",gsub("-unkown", "","ENSMMUG00000052585-unkown")), ]
all_lnc_1mb <- lapply(1:(length(lnc_ranges_complete_de)), function(index) unlist(get_n_closest_gr(lnc_ranges_complete_de[index], mrnas_ranges, n = 15, maxgap = 1000000L))[c(rep(FALSE,1), TRUE)]$gene_name)
do_go(unlist(all_lnc_1mb))

pairs_list <- lapply(1:(length(lnc_ranges_complete_de)), function(index) (get_n_closest_gr(lnc_ranges_complete_de[index], mrnas_ranges, n = Inf, maxgap = 1000000L)))[[1]]

myeloids <- subset(immune.combined, idents = "Myeloid")
count_matrix <- myeloids@assays$RNA@counts
correlations <- lapply(pairs_list, function(x) calc_correlation(x, count_matrix, type = "spearman"))
pvals <- (lapply(correlations, function(correlation) ifelse(length(correlation)== 1, print(1),print(correlation$p.value))))
rhos <- as.vector(lapply(correlations, function(correlation) ifelse(length(correlation)== 1, print(NA),print(correlation$estimate))))
# Select the significant one
rhos[pvals < 0.1]
pvals[pvals < 0.1]
correlated <- unlist(pairs_list[pvals < 0.1])[c(rep(FALSE,1), TRUE)]$gene_name

do_go(unlist(correlated))

log(1.2)
df <- data.frame(a = c(1,2), c = c("THY1", "ENSMMUG00000061948"))
ggplot(df, aes(x = a, fill = c ))+geom_bar()+scale_fill_manual(values  = c("navy", "#870052"))
```



# -------------------------------------------
# -------------------------------------------
# Analysis colocation in genomic regions

```{r colocation}
# 1) For each gene lncrna find neighbour protein coding genes. ( either overlapping or in close region) --> putative pairs
# 2) Filter: Check the pearson correlation of the identified pairs, reatain the ones w/ |pearson correlation coefficient| > 0.6
# ==> Only if something nice comes out 
# 3) Validation: Simulate random pairs 
# 4) Validation: Compare the correlation of the ones identified with the random ones. 
library(zeallot)
# --------------------------------------------------------------
#           Get 5 closest genes to each gene of interest
# --------------------------------------------------------------
# Find pairs under maximum overlap
get_n_closest_gr <- function(gr_gene, gr_hits =mrnas_ranges , n = 5, maxgap =1000000L ){
  p <- findOverlapPairs(gr_gene, get_gene_only(gr_hits), maxgap=maxgap, ignore.strand = TRUE)
  pairs_list <- zipup(p)
  print(gr_gene$gene_name)
  # Compute thedistances and select the first 5
  distances <- unlist(lapply(pairs_list, function(pair) GenomicRanges::distance(pair[1], pair[2], ignore.strand = TRUE) ))
  if(!is.null(distances)){
    n <- min(length(pairs_list), n)
    closest_5_pairs <-  pairs_list[order(distances)][1:n]
  }else{
    return(NULL)
  }
  return(closest_5_pairs)
}


get_closest_genes <- function(genes, original_range, hit_range, n = 10, maxgap = 100000L ){
  original_range <- original_range[original_range$gene_id %in% gsub("-", "_",gsub("-unkown", "",genes)), ]
  ranges_5_cl_genes <- lapply(1:(length(original_range)), function(index) unlist(get_n_closest_gr(original_range[index], hit_range, n = n, maxgap = maxgap))[c(rep(FALSE,1), TRUE)]$gene_name)
return(ranges_5_cl_genes)
}


# prepare ranges
mrnas_ranges <- ref[ref$gene_name %in% mrna_ref_id]
lnc_ranges <- ref[ref$gene_id %in% gsub("-", "_",gsub("-unkown", "",all_lncrnas)),]
# add gene entries for novel lncrna, which only present transcripts and exons
genes_lnc_ranges <- get_gene_only(lnc_ranges)$gene_id
missing_gene_lines <- unique(lnc_ranges$gene_id)[!(unique(lnc_ranges$gene_id) %in% genes_lnc_ranges)]
gr_list <- lapply(missing_gene_lines, function(x) (get_gene_range(x, lnc_ranges)))
missing_gene_ranges <- do.call(base::c,gr_list)
lnc_ranges_genes <-c(get_gene_only(lnc_ranges), missing_gene_ranges)


# Group DE lnc 
lnc_myeloid_late <- get_de_names(mast_myeloid_late, subset = all_lncrnas, FCTHRESHOLD = log(1.0))
lnc_myeloid_middle <- get_de_names(mast_myeloid_middle, subset = all_lncrnas, FCTHRESHOLD = log(1.0))
lnc_myeloid_early <- get_de_names(mast_myeloid_early, subset = all_lncrnas, FCTHRESHOLD = log(1.0))
de_myeloid <- c(lnc_myeloid_late, lnc_myeloid_middle, lnc_myeloid_early)

lnc_T_late <- get_de_names(mast_T_late, subset = all_lncrnas, FCTHRESHOLD = log(1.2))
lnc_T_middle <- get_de_names(mast_T_middle, subset = all_lncrnas, FCTHRESHOLD = log(1.2))
lnc_T_early <- get_de_names(mast_T_early, subset = all_lncrnas, FCTHRESHOLD = log(1.2))
de_t <- c(lnc_T_late, lnc_T_middle, lnc_T_early)

lnc_B_late <- get_de_names(mast_B_late, subset = all_lncrnas)
lnc_B_middle <- get_de_names(mast_B_middle, subset = all_lncrnas)
lnc_B_early <- get_de_names(mast_B_early, subset = all_lncrnas)
de_b<- c(lnc_B_late, lnc_B_middle, lnc_B_early)

de_all <- c(de_myeloid, de_t,de_b)
closest_m <- get_closest_genes(unique(de_myeloid), lnc_ranges_genes, get_gene_only(mrnas_ranges), Inf, 1000000L)

closest_t <- get_closest_genes(unique(de_t), lnc_ranges_genes, get_gene_only(mrnas_ranges), Inf, 1000000L)

closest_b <- get_closest_genes(unique(de_b), lnc_ranges_genes, get_gene_only(mrnas_ranges), Inf, 1000000L)

closest_myeloid <- get_closest_genes(unique(de_myeloid), lnc_ranges_genes, get_gene_only(mrnas_ranges), Inf, 1000000L)

closest_all <- get_closest_genes(unique(de_all), lnc_ranges_genes, get_gene_only(mrnas_ranges), Inf, 1000000L)


do_go(unique(unlist(closest_all)))
do_go(unique(unlist(closest_t)))
do_go(unique(unlist(closest_b)))
do_go(unique(unlist(closest_myeloid)))
do_go(unique(unlist(closest_m)))


# Extract mrnas from significant correlations 
do_go <- function(list, keytype = "SYMBOL"){
  ego <- clusterProfiler::enrichGO(gene = list,OrgDb =  org.Hs.eg.db, keyType = keytype,ont = "BP")
  barplot(ego, showCategory=10)
  clusterProfiler::dotplot(ego)
}

```


# -------------------------------------------
# -------------------------------------------
# Analysis cis-action/coregulation 

```{r cis-regulation}
# 1) For each gene lncrna find neighbour protein coding genes. ( either overlapping or in close region) --> putative pairs
# 2) Filter: Check the pearson correlation of the identified pairs, reatain the ones w/ |pearson correlation coefficient| > 0.6
# ==> Only if something nice comes out 
# 3) Validation: Simulate random pairs 
# 4) Validation: Compare the correlation of the ones identified with the random ones. 
library(zeallot)

# 1. extract the count matrix
count_matrix <- log2(myeloids@assays$RNA@counts+.5)
count_matrix <- myeloids@assays$RNA@counts
rownames(count_matrix) <- gsub("-unkown","",rownames(count_matrix))
exp_matrix <-myeloids@assays$integrated@data

# 2. 
mrnas_ranges <- ref[ref$gene_name %in% mrna_ref_id  ]
lnc_ranges <- ref[ref$gene_id %in% gsub("-", "_",gsub("-unkown", "",all_lncrnas))]


lnc_ranges
# Add missing gene entries 
genes_lnc_ranges <- get_gene_only(lnc_ranges)$gene_id
missing_gene_lines <- unique(lnc_ranges$gene_id)[!(unique(lnc_ranges$gene_id) %in% genes_lnc_ranges)]
gr_list <- lapply(missing_gene_lines, function(x) (get_gene_range(x, lnc_ranges)))
missing_gene_ranges <- do.call(base::c,gr_list)
c(gr_list[1][[1]],gr_list[2][[1]])

lnc_ranges_complete <-c(get_gene_only(lnc_ranges), missing_gene_ranges)

# Only select the DE interestin

lnc_ranges_complete_de <- lnc_ranges_complete[lnc_ranges_complete$gene_id %in% gsub("-", "_",gsub("-unkown", "",de_lnc_myelid_late_ordered)), ]

# ----------------------
# Find pairs allowing 10K gap on both sides

pairs <- findOverlapPairs(lnc_ranges_complete_de, get_gene_only(mrnas_ranges), maxgap=1000000L, ignore.strand = TRUE)
# Get a list of ranges objects in pairs
pairs_list <- zipup(pairs)
length(pairs_list)

# How many lncrnas in total vs how many are in pairs
length(unique(lnc_ranges$gene_id)); length(unique(mrnas_ranges$gene_id)); length(pairs)

# get the counts of the interesting genes
calc_correlation <- function(pair, count_matrix, type = "pearson"){
  gene1 <- str_replace_all(pair[1]$gene_id, "_" , "-")
  gene2 <- str_replace_all(pair[2]$gene_id, "_" , "-")
  print(gene1)
  print(gene2)
  c1 <- count_matrix[rownames(count_matrix)== gene1,]
  c2 <- count_matrix[rownames(count_matrix)== gene2,]
  print("----")
  # calculate the correlation coefficient 
  if(length(c1)> 0 && length(c2)>0){ 
    #pc <- bayes.cor.test(c1, c2) 
    pc <- cor.test(c1, c2, method = c(type)) 
  }else{ pc = cor.test(c(1:100),c(1:100), method = c(type))  }
  return(pc)
}
# Calc correlation values

correlations <- lapply(pairs_list, function(x) calc_correlation(x, count_matrix, type = "spearman"))
pvals <- (lapply(correlations, function(correlation) print(correlation$p.value)))
rhos <- as.vector(lapply(correlations, function(correlation) print(correlation$estimate)))

# Obtain significant correlations
significant_correlations <- correlations[pvals < 0.1 & rhos !=1]
significant_pairs <- pairs_list[pvals < 0.1 & rhos !=1]
significant_correlations

# Extract mrnas from significant correlations 
do_go <- function(list, keytype = "SYMBOL"){
  ego <- clusterProfiler::enrichGO(gene = list,OrgDb =  org.Mmu.eg.db, keyType = keytype,ont = "BP",universe =unique(ref$gene_name))
  barplot(ego, showCategory=10)
  clusterProfiler::dotplot(ego)
}
correlated_mrnas <-unlist(significant_pairs)[c(rep(FALSE,1), TRUE)]$gene_id
do_go(correlated_mrnas, keytype = "ENSEMBL")

keytypes(org.Mmu.eg.db)

```